Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-06-01 paper/main.tex
xi and operator figs

\IEEEPARstart{V}isualizations are expected to be a faithful translation of the inherent structure of the data being visualized; therefore the components of visualization software libraries that translate data to graphics are also expected to preserve this structure. Motivated by a need to develop new library components that adapt to modern data needs, especially high dimensional multivariate and distributed data, we propose a framework that provides a language to express that topological and field structure of the input data that must be mapped to the visualization and which changes to data must be expressed in visualization, and a functional framework that describes composability in terms of functional algebraic operations. This allows complex visualizations to be built from simple verifiable parts\cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}, which makes these components easier to drop into the existing library and be used by domain specific library developers to build new libraries that work with existing ones.  While many of these elements have been partially described elsewhere, our framework naturally combines them in a unified framework based on ideas in category theory. Further we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib. In this paper, we introduce a mathematical framework for describing, constructing, and testing visualization library components and provide a toy example illustrating these concepts.

\section{Related Work}
In this work, we propose a uniform method for expressing the structure that visualizations are expected to preserved by combining and generalizing previous work that describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent changes to data and graphics. We propose that a unified abstraction model can guide developers in building structure aware composable, reusable software library components with more consistent interfaces.

@@

The locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space. The locality axiom asserts that two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\desection^{a}|{\openset} =  \dsection^{b}|_{openset}$. The gluing axiom asserts that the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \sqcup \openset_{j}} = \dsection^{i}|_{\openset_i} \sqcup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$ The gluing axiom generalizes to expresses an equivalency between a concatenated representation of data and distributed representation of the same data. In \autoref{fig:atct:networkbundle} this axiom means that the sections over each node and the edge can be combined into one section over the piece of the network in the figure.

@@

\subsection{Data and Graphic Correspondence}
\label{sec:atct:xi}
A property of sheaves is that, given two sheaves, if there exists a map between the base spaces of each sheave, then we can construct transport functors that can move each sheaves over the other base space. This allows us to move data over the corresponding graphic base space and graphics over the corresponding data base space. The data sections defined in \autoref{eq:atct:fb} are sections of sheaves, introduced in \autoref{sec:atct:sheaves}, over open sets $\openset \subseteq \dbase$ in a base space \dbase\ that encode the topology of the data topology. The graphic sections, defined in \autoref{eq:atct:fb_graphic_section}, are sections of sheaves over open sets $\opensetg \subseteq \gbase$ in a base space \gbase\ that encodes the topology of the graphic. We propose that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc 
\begin{equation}
  \label{eq:atct:xi}
  \vindexc: \opensetgc \rightarrow \opensetc 
\end{equation}
such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$.

@@

 The pushforward $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
 \begin{equation}  
  \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} \in \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} 
\end{equation}
such that it provides a way to look up which graphic  corresponds with a data index
\begin{equation}
  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)} = \gsectionc(\gbasepointc)\;\forall \gbasepointc\in \vindexprec(\dbasepointc)
\end{equation}

>>>>

\IEEEPARstart{V}isualizations are expected to be a faithful translation of the inherent structure of the data being visualized; therefore the components of visualization software libraries that translate data to graphics are also expected to preserve this structure. Motivated by a need to develop new library components that adapt to modern data needs, especially high dimensional multivariate and distributed data, we propose a framework that topological and field structure of the input data that must be mapped to the visualization and which changes to data must be expressed in visualization, and a a functional design framework that describes composability in terms of functional algebraic operations. \note{this may be a good place to incorporate types} This allows complex visualizations to be built from simple verifiable parts\cite{huHowFunctionalProgramming2015,hughesWhyFunctionalProgramming1989}, which makes these components easier to drop into the existing library and be used by domain specific library developers to build new libraries that work with existing ones.  While many of these elements have been partially described elsewhere, our framework naturally combines them in a unified framework based on ideas in category theory. Further we see the application of these ideas in low level visualization library design as our motivation, and we hope these ideas will help shape the reorganization of critical data visualization libraries, such as Matplotlib. The contribution of this paper is a uniform method for describing the structure in data and a mathematical framework for describing, constructing, and testing structure preserving visualization library components.

\section{Related Work}
We build on the exentsive work codifying the structure of visualization data to develop a more generalizable method for expressing the structure that visualizations are expected to preserved. Broadly, previous work describes structure as a combination of the topological properties of the data, the mathematical structure of the data fields, and equivalent changes to data and graphics. We propose that a unified abstraction model can guide developers in building structure aware composable, reusable software library components with more consistent interfaces.

@@

The locality and gluing axioms \cite{bakerMathsSheaf} of a sheaf describe the conditions for verifying that the sections over subspaces are part of a cohesive section over the whole space. The locality axiom asserts that two sections in a sheaf are equal $\dsection^{a} = \dsection^{b}$ when they evaluate to the same values over the same set of open sets $\dsection^{a}|{\openset} =  \dsection^{b}|_{\openset}$. The gluing axiom asserts that the union of sections defined on specific open sets is equivalent to one big section over the union of spaces $\dsection|_{\openset_{i} \cup \openset_{j}} = \dsection^{i}|_{\openset_i} \cup \dsection^{j}|_{\openset_j}$ if these sections agree on overlaps $\dsection^{i}|_{\openset_i\cap\openset_j} =  \dsection^{j}|_{\openset_i\cap\openset_j}$ The gluing axiom generalizes to expresses an equivalency between a concatenated representation of data and distributed representation of the same data. In \autoref{fig:atct:networkbundle} this axiom means that the sections over each node and the edge can be combined into one section over the piece of the network in the figure.

@@

\subsection{Data and Graphic Correspondence}
\label{sec:atct:xi}
A property of sheaves is that, given two sheaves, if there exists a map between the base spaces of each sheave, then we can construct transport functors that can move each sheaves over the other base space. This allows us to move data over the corresponding graphic base space and graphics over the corresponding data base space. The data sections defined in \autoref{eq:atct:fb} are sections of sheaves, introduced in \autoref{sec:atct:sheaves}, over open sets $\openset \subseteq \dbase$ in a base space \dbase\ that encode the topology of the data topology. The graphic sections, defined in \autoref{eq:atct:fb_graphic_section}, are sections of sheaves over open sets $\opensetg \subseteq \gbase$ in a base space \gbase\ that encodes the topology of the graphic. We propose that there exists a continuous \textcolor{functor}{map between topological spaces} \vindexc 
\begin{equation}
  \label{eq:atct:xi}
  \vindexc: \opensetgc \rightarrow \opensetc 
\end{equation}
such that every $\opensetg \subseteq \gbase$ must map to a corresponding set $\openset \subset \dbase$. The functor $\vindex$ is surjective such that for every point $\dbasepoint \in \openset$ there is a corresponding set of points $\{\gbasepoint | \gbasepoint \in  \vindexpre(k)\}$ for all $\gbasepoint in \opensetg$.

@@

 The pushforward $\vindexpushc$ transports sheaves of sections on $\opensetgc$ over $\openset$
 \begin{equation}  
  \vindexpushc\gsectionc: \opensetc \rightarrow \vindexpushc \gtotalc\restriction_{\opensetc} \in \cgamma{\opensetc}{\vindexpushc\gtotalc\restriction_{\opensetc}} 
\end{equation}
such that it provides a way to look up which graphic  corresponds with a data index
\begin{equation}
  \vindexpushc\gsectionc(\dbasepointc) = \gsectionc\restriction_{\vindexprec(\dbasepointc)} = \{\gsectionc(\gbasepointc)\;for\;all\; \gbasepointc\in \vindexprec(\dbasepointc)\}
\end{equation}
